\documentclass{article}

\usepackage{fullpage}
\usepackage{amsmath,amsfonts,amsthm,amssymb}

\begin{document}
\title{CS 457, Data Structures and Algorithms I\\
Fifth Problem Set}
\date{November 20, 2019}
\maketitle
\begin{center}
\textbf{Due on December 2. Collaboration is not allowed. Contact Daniel and me for questions.}
\end{center}

For any algorithms that you propose, also provide a (high level) justification regarding their worst-case running time bounds, and a brief explanation regarding their correctness.
\begin{enumerate}
\item (25 pts) Consider a variation of the $\textsc{Strongly-Connected-Components}(G)$ (see Page 617 of your textbook) which, rather than calling DFS($G^T$) in Step 3, calls DFS($G$) instead (everything else remains the same). Provide a graph $G=(V, E)$ for which this variation is incorrect, i.e., it does not output the correct set of strongly connected components of $G$. For full credit, explain what the variation of the algorithm would do for this graph $G$: 
\begin{itemize}
\item[i)] provide $v.d$ and $v.f$ for all $v\in V$ for Step 1 (you can choose the initial ordering of the vertices),
\item[ii)] provide $v.d$ and $v.f$ for all $v\in V$ for Step 3, 
\item[iii)] provide the set of connected components that it would output, and 
\item[iv)] explain why this output is incorrect.
\end{itemize}

\item (25 pts) Provide an algorithm that, given a directed acyclic graph $G=(V, E)$ and two vertices $s, t\in V$, returns the number of simple paths from $s$ to $t$ in $G$. Your algorithm's worst-case running time should be $O(m +n)$. 


\item (25 pts) An Eulerian tour of a strongly connected directed graph $G=(V, E)$ is a directed cycle that contains {\em every} edge in $E$ exactly once (at least once and no more than once). First, prove that $G$ has an Eulerian tour {\em if and only if} every vertex in $V$ has its in-degree equal to its out-degree. Then, using this fact, provide a $O(m)$ algorithm that returns an Eulerian tour of $G$, if one exists, or reports that no Eulerian tour exists. Hint: your algorithm can merge edge-disjoint cycles.

\item (25 pts) Provide a DFS-based algorithm that takes as input a graph $G=(V,E)$ and determines if there exists an edge $e\in E$ such that removing $e$ from $E$ increases the number of connected components of $G$. If such an edge exists, your algorithm should return all of the edges in $E$ with this property. For instance, if $V=\{v_1, v_2, v_3\}$ and $E=\{(v_1, v_2), (v_2, v_3)\}$, then $G=(V,E)$ has a single connected component. Removing either one of the two edges in $E$ would increase the connected components from one to two, so the algorithm should return both of these edges in this example. The worst-case running time of the algorithm should be $O(m+n)$.

\end{enumerate}
\end{document}