\documentclass{article}

\usepackage{fullpage}
\usepackage{amsmath,amsfonts,amsthm,amssymb}

\begin{document}
\title{CS 457, Data Structures and Algorithms I\\
Sixth Problem Set}
\date{December 2, 2019}
\maketitle
\begin{center}
\textbf{Due on December 6 at 11:59 P.M. Collaboration is not allowed. Contact Daniel and me for questions.}
\end{center}

For full credit, for any algorithms that you propose, also provide a (high level) justification regarding their worst-case running time bounds, and a brief explanation regarding their correctness. Throughout the problem set $n$ denotes the number of vertices of a graph and $m$ denotes the number of edges.
\begin{enumerate}

\item (20 pts) Your textbook defines the \emph{connected components} of undirected graphs and the \emph{strongly connected components} of directed graphs (see pages 1170-1171). Adding edges to a graph increases the connectivity and may lead to a decrease in the number of such components. i) If you add an edge to an undirected graph $G$, what is the largest possible drop in the number of connected components of $G$ that this can possibly lead to? Similarly, ii) if $G$ is a directed graph and you add a directed edge to it, what is the largest possible drop in the number of strongly connected components of $G$ that this can lead to? Provide both upper and lower bound arguments to support your claims. For instance, if you claim that the largest possible drop is $n/4$, then provide an example where the drop is $n/4$ and an argument why it can be no more than $n/4$ in general. 


\item (25 pts) You are given a weighted graph $G=(V, E)$ as well as a minimum spanning tree $T$ of $G$. Then, a new edge $e$ is added to $G$ and you want to compute a minimum spanning tree for the new graph $G'=(V, E\cup\{e\})$. Provide an algorithm $\textsc{New-MST}(G,T,e)$ that computes a minimum spanning tree $T'$ for $G'$ in time $O(n)$. Clearly, $T'$ could be computed from scratch using Prim's or Kruskal's algorithm, or one could use algorithms covered in recitation, but all of these solutions would require time $\omega(n)$ which is not good enough.

\item (25 pts) You are given a directed graph $G=(V,E)$ along with a function $w:E\to \mathbb{R_+}$ that assigns positive weights to all the edges. This graph may contain directed cycles and your goal is to check if such cycles exist in $G$ and, if they do, to return one with the smallest possible weight. The running time of your algorithm should be $O(n^3)$ or $O(n^2m)$ (whichever one you prefer). You can use any algorithm from Chapter 24 as a subroutine.


\end{enumerate}
\end{document}